---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bopR

<!-- badges: start -->
[![R-CMD-check](https://github.com/sciensanogit/bopR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sciensanogit/bopR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

bopR is the Belgian open population repository for R

## Installation

You can install the development version of bopR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sciensanogit/bopR")
```

## Population datasets

`bopR` stands for Belgian open population repository, as it makes Belgian population data available in R. The `bopR` package contains population files on different geographical levels (municipalities - districts - provinces -  regions - communities - country). Herefore, `bopR` includes datasets on the structure of the population based on the national registries, and datasets on demographic projections. Herefore, `bopR` relies on the open data that is available at Statistics Belgium (Statbel) and the Federal Plan Bureau. In addition, the `bopR` package also includes midpopulation data as calculated in the framework of the Standard Population Mortality Analysis (SPMA), as developed by Sciensano.

### Statbel

Datasets from Statbel are derived from the Statbel's OPEN DATA repositories. In the `bopR` package they are called as follows:

* *BE_POP_MUNTY*: municipalities
* *BE_POP_ARRD*: districts
* *BE_POP_PROV*: provinces (+ Brussels Capital Region for simplicity)
* *BE_POP_RGN*: regions
* *BE_POP_COMTY*: communities
* *BE_POP*: country

```{r, echo = FALSE}
library(ggplot2)
data("BE_POP")
dta <- aggregate(
  data = subset(BE_POP, subset = SEX == "MF" & AGE == "ALL"),
  POPULATION ~ YEAR,
  FUN = sum)

ggplot(data = dta) +
  geom_line(aes(x = as.numeric(YEAR), y = POPULATION)) +
  ylim(0,14000000) +
  theme_bw()
```

### Federal Planning bureau

Datasets from Statbel are derived from the Federal Planning Bureau repositories. In the `bopR` package they are called as follows:

* *BE_POP_PROJ_MUNTY*: municipalities
* *BE_POP_PROJ_ARRD*: districts
* *BE_POP_PROJ_PROV*: provinces (+ Brussels Capital Region for simplicity)
* *BE_POP_PROJ_RGN*: regions
* *BE_POP_PROJ_COMTY*: communities
* *BE_POP_PROJ*: country

```{r, echo = FALSE}
library(ggplot2)
data("BE_POP_PROJ")
dta <- aggregate(
  data = subset(BE_POP_PROJ, subset = SEX == "MF" & AGE == "ALL"),
  POPULATION ~ YEAR,
  FUN = sum)

ggplot(data = dta) +
  geom_line(aes(x = as.numeric(YEAR), y = POPULATION)) +
  ylim(0,14000000) +
  theme_bw()
```


for more information on the content of each dataset, you can always consult the help-file by `?BE_POP`. The population structure of Belgium over the years was as follows:

```{r, echo = FALSE}
# data("BE_POP")
# betab <- aggregate(
#   data = subset(BE_POP, subset = SEX == "MF" & AGE == "ALL"),
#   POPULATION ~ YEAR,
#   FUN = sum)
# 
# kableExtra::kable(betab)
```

## Population data by statistical sector

the `bopR` package also contains population files on statistical sector level (not by sex or age). The dataset is named: `BE_POP_SECTOR`.

## Population projection datasets

the `bopR` package also contains population projection files on different geographical levels (districts - provinces -  regions - country). The datasets are named as follows:

## Data sources

The idea for this package is inspired by the [BelgiumMaps.StatBel package](https://github.com/bnosac/BelgiumMaps.StatBel), developed by Jan Wijffels at [BNOSAC](http://www.bnosac.be/index.php/blog/55-belgiummaps-statbel-r-package-with-administrative-boundaries-of-belgium).

The population files are obtained from the Statbel open data and the open data from the federal plan bureau (FPB) of Belgium:

* **https://statbel.fgov.be/en/themes/population** includes information on the population of Belgium on the 1st of January for each year.
* **https://www.plan.be/publications/publication-2255-nl-demografische_vooruitzichten_2021_2070_update_sterke_opwaartse_herziening_van_de_bevolkingsgroei_in_2022** includes information on the population projections of Belgium.

To calculate the mid-year population of year `x`, one should average the population of year `x` and year `x+1` .

